Towards a Machine Learning approach for Aggregate Computing

### 2021-08-11 paper-2021-woa-ac-ml.tex
Some characteristics of AC in ML-perspective

%% The first command in your LaTeX source must be the \documentclass command.
%%
%% Options:
%% twocolumn : Two column layout.
%% hf: enable header and footer.
\documentclass[
% twocolumn,
% hf,
]{ceurart}
\usepackage[colorinlistoftodos,prependcaption]{todonotes}
\usepackage{xcolor}
%%
%% One can fix some overfulls
% \sloppy

@@

>>>>

%% The first command in your LaTeX source must be the \documentclass command.
%%
%% Options:
%% twocolumn : Two column layout.
%% hf: enable header and footer.
\documentclass[
twocolumn,
% hf,
]{ceurart}
\usepackage[colorinlistoftodos,prependcaption]{todonotes}
\usepackage{xcolor}
%%
%% One can fix some overfulls
\sloppy

\newcommand{\meta}[1]{{\color{blue}#1}}

@@

In this section,
 we analyse the characteristics of the aggregate computing paradigm,
 map these to relevant machine learning contributions in literature,
 and summarise the analysis into a roadmap.

\paragraph{Multi- and many-agent system}
%
An aggregate system is a multi-agent system
 and, often, a \emph{many}-agent system
 where a large number (hundreds or more)
 of autonomous entities are programmed to achieve 
 some collective behaviour by means of \emph{repeated} 
 sensing, computation, communication, and actuation steps.
%
\meta{TODO present approaches for learning in MAS, e.g., through prominent surveys}

\paragraph{Neighbour-based or indirect (environment-mediated) interaction}
%
In aggregate systems,
 a device can only directly interact with neighbours.
%
Data flows may be implemented
 to support indirect communication across multiple hops;
 however, information from devices far away tends to be obsolete.
%
The environment can also be used, via stigmergy,
 for indirect communication;
 however, a device can typically only access 
  its very local portion of the environment.
%
In other words, the system state is only \emph{partially observable}.
%
\meta{What ML in these kinds of systems?}

\paragraph{Learning goal: collective behaviour}
%
The devices of an aggregate 
 must cooperatively learn the ``aggregate program'',
 namely the collective behaviour 
 that achieves a particular \emph{global goal}
 in a decentralised, resilient way.
%

\paragraph{Self-organisation, swarm robotics}
%
\meta{TODO}

\paragraph{Emergent behaviour}
%
The collective behaviour of an aggregate system
 \emph{emerges} 
 from a dense network of computations and interactions
 in an evolving environment.
%

\paragraph{Delayed, global rewards}
%
In the light of the above properties,
 if we consider a learning approach based on rewards,
 we must observe that such rewards would be
 delayed and mostly global---i.e.,
 it may not be easy to solve the credit-assignment problem
 to distinguish good from bad individual contributions
 to the global outcome.

\paragraph{Transient phase}
%
When an aggregate behaviour is carried out,
 we often focus on the stabilised outcomes,
 tolerating outcomes that are being generated
 during the transient phase.
%
However, intermediate results are often important,
 and invariants should hold throughout the entire computation.

\paragraph{Dynamic topology}
%
If we admit \emph{mobility} and \emph{failure},
 then we must also consider 
 the possibility of changes in neighbourhoods
 and hence the need of dealing with dynamic topologies.

\subsection{Roadmap}
